\documentclass{article}

\usepackage[hidelinks]{hyperref}
\usepackage[italian, main=english]{babel}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{microtype}

\lstset{
    language=Python,
    basicstyle=\ttfamily\small,
    keywordstyle=\color{blue}\bfseries,
    stringstyle=\color{green!50!black},
    commentstyle=\color{gray}\itshape,
    numberstyle=\tiny\color{gray},
    numbers=left,
    stepnumber=1,
    showstringspaces=false,
    breaklines=true,
    frame=single,
    captionpos=b,
    tabsize=4
}

\title{Internship Project - UniVR \break
Skill Gap Evaluation and Course Recommendations}
\author{Simone Pedretti}
\date{\today}

\begin{document}

\maketitle

\tableofcontents

\newpage

\section{Introduction}
This report details the work carried out during my internal internship at 
\foreignlanguage{italian}{\textit{Università degli Studi di Verona}}. 
The primary objective of this project was the design and development of a web-based 
platform tailored for skill management, competence tracking, and course recommendation.

From a technical perspective, the application follows a standard client-server 
architecture, with a clear separation of concerns between the frontend and the backend. 
The backend system, hosted on the server to handle the core business logic and data 
management, was developed in Python utilizing the \textbf{FastAPI} framework. 
It heavily relies on the \textbf{JSON} format for efficient data structuring 
and API communication. 
The frontend system, on the client side, was implemented using standard web 
technologies, specifically \textbf{HTML} and \textbf{CSS}. This ensures 
a clean, responsive, and accessible experience for both individual users and 
organizational profiles interacting with the system.

The following sections will provide a detailed overview of the specific features implemented for the 
different types of users, detailing the browsing capabilities, profile management, 
and the skill gap evaluation mechanisms. Furthermore, the report will explore 
the underlying technical components that power these functionalities: 
final chapters will discuss the integration of the \textbf{ESCO API} for standardized 
competence mapping, the strategies adopted for data management, the frontend design 
principles utilizing HTML and CSS, and finally, the testing procedures conducted to 
ensure the reliability and robustness of the system.

\newpage

\section{Individual User Features}
This section outlines the functionalities available to individual users interacting 
with the platform. Within this system, a user can hold one of two distinct roles: 
a standard Employee or a Manager. While an Employee utilizes the platform primarily 
for personal skill tracking and career development, a Manager is a user who has been 
elevated to this status by their affiliated organization. Managers inherit all standard 
employee capabilities but are additionally granted administrative privileges to 
oversee projects and coordinate team members within their company.

\subsection{Home}
The Home dashboard serves as the primary entry point for users, offering intuitive 
search and management tools depending on their assigned role.

\subsubsection{Browsing for Roles and Skills}
Users can explore various professional profiles through the role browsing feature. 
This search mechanism integrates the ESCO API, ensuring that role definitions, 
responsibilities, and terminologies are standardized and perfectly aligned with 
the European labor market classifications.

Similarly to the role search, users can browse for individual competencies. 
By querying the ESCO API, the system retrieves a list of macthing skills. 
This allows users to accurately identify specific 
abilities.

\subsubsection{Project Manager}
This dedicated area is exclusively accessible to users with Manager privileges. 
Here, managers can create, organize, and oversee various internal projects. 
They have the authority to assign team members to these projects, provided those 
individuals belong to the same organization. Furthermore, managers can select specific 
target roles for a project to calculate Cedefop-based employment forecasts and 
generate preliminary course recommendations for the team. A more comprehensive 
explanation of the forecasting and recommendation mechanisms is detailed in the 
Profile section below.

\subsection{Profile}
The Profile section acts as the personal hub for the user, centralizing account 
settings, competence tracking, and career development tools.

\subsubsection{Target Roles and Skills Management}
Both employees and managers can access this area to view their personal data and 
securely change their passwords. More importantly, users can manage their professional 
portfolio by viewing, adding, or removing their personal skills and desired target roles. 
If a user is currently affiliated with an organization, a dedicated subsection 
displays their ``organizational skills''---competencies that are directly assigned 
and managed by the company. Through this interface, users also retain the autonomy 
to officially leave their current organization if needed.

\subsubsection{Cedefop Forecast and Skill Gap Evaluation}
Once a user has defined their current skill set and selected one or more target roles, 
the platform can perform a comprehensive skill gap analysis. The system compares the 
user's existing competencies against the specific requirements of the desired roles,
highlighting areas for improvement. Additionally, the application retrieves Cedefop 
forecast data related to these target roles, offering the user valuable insights 
into future labor market trends, job demand, and employment projections.

\subsubsection{Course Recommendation}
Based on the exact skill gaps identified in the previous step, the platform provides 
customized educational suggestions. This recommendation engine---which is the same core 
logic utilized in the Project Manager dashboard---analyzes the missing competencies 
and cross-references them with available training resources. By doing so, the system 
generates a highly personalized learning path, proposing the most relevant courses to 
help the user efficiently bridge the gap between their current capabilities and their 
professional goals.

\newpage

\section{Organization Features}
This section details the functionalities available to Organization accounts. 
Unlike individual users who focus on personal career development, an Organization 
profile represents the company as a whole. Its primary purpose is to oversee the 
workforce, manage internal training resources, and analyze competence gaps at a macro 
level, providing an administrative overview of the company's human resources.

\subsection{Home}
The Home dashboard for an organization is designed to provide a high-level summary 
of ongoing activities, focusing primarily on project monitoring.

\subsubsection{Projects}
Through this interface, the organization can access a comprehensive list of all 
active projects currently overseen by its internal Managers. It is important to note 
that this is a strictly read-only view: the organization can monitor the status, but 
the direct management, creation, and modification of the projects remain the sole 
responsibility of the delegated Managers.

\subsection{Profile}
The Profile area serves as the central administrative hub for the organization, 
allowing the management of account credentials, personnel, and proprietary training 
materials. As with individual users, this section also includes basic settings to 
view corporate data and securely update the account password.

\subsubsection{Team Members}
The organization can view its current workforce between employees and managers. 
To expand the team, the organization can send an official invitation by 
entering a specific username into the recruitment form, facilitated by a 
list of all registered platform users. The invited user then retains the autonomy 
to accept or decline the request. 
Furthermore, the organization holds the authority to adjust user privileges within 
the company, allowing them to promote standard employees to the Manager role or demote 
them as organizational needs dictate.

\subsubsection{Global Skill Gap Evaluation}
Provided the organization has active projects, it can perform a macroscopic 
\mbox{analysis} of its workforce capabilities. Instead of evaluating a single 
user, this feature calculates a global skill gap by comparing the aggregated 
competencies of the current employees against the collective requirements of 
the ongoing projects. Based on these company-wide deficiencies, the platform suggests 
HR-oriented courses. These specific recommendations are designed for strategic 
corporate upskilling and human resource planning, rather than individual learning paths.

\subsubsection{Courses Management}
The platform allows organizations to act as educational providers by creating 
and distributing their own training modules. Through this interface, the company 
can view a complete catalog of its authored courses, seamlessly create new ones, 
and edit or delete existing materials. A critical feature of this management tool 
is the visibility toggle: courses can be published as ``public'' to be accessible 
by any user on the platform, or set to ``private,'' ensuring that only the 
organization's officially affiliated employees can view and benefit from them.

\section{Guest Features}
The platform also accommodates unregistered users by providing a restricted ``Guest'' 
access mode. For individuals who prefer not to create a formal account, the system 
offers a streamlined interface to explore the core competency catalog. Through this 
guest portal, users can perform basic searches for professional roles and individual 
skills, which are directly powered by the ESCO API. 

However, compared to registered employees or organizations, this browsing experience 
is intentionally synthesized and limited in scope. Guests can view roles 
of the queried terms, but they cannot access 
advanced, stateful features such as profile creation, skill gap evaluation, or 
personalized course recommendations. This approach allows external users to evaluate 
the platform's core search capabilities while maintaining data privacy and 
encouraging formal registration for full access.

\newpage

\section{ESCO API Integration}
A core component of the platform is the integration with the ESCO API, hosted by 
the European Commission. The system interacts with this external service using the 
Python \texttt{requests} library to fetch standardized data regarding professional 
roles and specific skills.

The backend implements dedicated functions to query the API for single role 
details (including the ISCO classification code and essential skills), 
as well as search functions to retrieve lists of occupations or skills based 
on user input. For instance, the system requests the \texttt{hasEssentialSkill} 
relation from the API to automatically map the necessary competencies for a given 
job title. The JSON responses provided by the API are subsequently parsed and 
converted into internal \texttt{Role} and \texttt{Skill} data models to be utilized 
by the application's business logic.

\section{Data Management}
Given the platform's reliance on extensive datasets, particularly for the Cedefop 
employment forecasts, efficient data management was a priority. The original Cedefop 
data, initially provided in Excel format, was converted into structured JSON files. 

To ensure high performance and minimize read operations from the disk during 
execution, the application utilizes the FastAPI \texttt{app.state} object. 
Upon server startup, the system loads these JSON files directly into the server's RAM. 
This in-memory caching strategy significantly reduces the latency when users calculate 
skill gaps and request forecast data, allowing the application to respond to these 
computationally heavy requests almost instantaneously. User and organization data, 
as well as active invitations, are also managed via JSON-based local databases, 
allowing for a lightweight and easily deployable architecture.

\section{Testing and Quality Assurance}
To guarantee the stability and reliability of both the backend calculations and the 
frontend endpoints, a comprehensive testing suite was implemented using the 
\textbf{pytest} framework. The testing strategy covers HTML route rendering, 
internal backend logic (such as the skill gap evaluation algorithms), and proper 
HTTP status code responses.

To simulate client interactions without requiring a live server, the FastAPI 
\texttt{TestClient} is utilized. Furthermore, to isolate the testing environment 
and prevent the corruption of actual production data, the system employs the 
\mbox{\texttt{monkeypatch}} utility provided by pytest. This allows the automatic 
generation of temporary, isolated JSON directories for users, organizations, 
and invitations during the testing phase, which are automatically 
discarded upon completion. 

The following snippet illustrates the configuration of the unified mocking database 
used across the test suite:

\begin{lstlisting}[caption={Pytest configuration with TestClient and temporary mock databases}]
import pytest
from fastapi.testclient import TestClient
from app.main import app
from app.crud import crud_user, crud_org

@pytest.fixture(scope="module")
def client():
    with TestClient(app) as c:
        yield c

@pytest.fixture(autouse=True)
def mock_database(tmp_path, monkeypatch):
    # Create temporary directories
    temp_users_dir = tmp_path / "test_users"
    temp_orgs_dir = tmp_path / "test_orgs"
    temp_inv_dir = tmp_path / "test_invitations"
    
    temp_users_dir.mkdir(exist_ok=True)
    temp_orgs_dir.mkdir(exist_ok=True)
    temp_inv_dir.mkdir(exist_ok=True)

    # Forcing code to use new tmp dirs
    monkeypatch.setattr(crud_user, "DATA_DIR_USERS", str(temp_users_dir))
    monkeypatch.setattr(crud_user, "DATA_INV_DIR", str(temp_inv_dir))

    monkeypatch.setattr(crud_org, "DATA_DIR_ORGS", str(temp_orgs_dir))
    monkeypatch.setattr(crud_org, "DATA_INV_DIR", str(temp_inv_dir))

    yield
\end{lstlisting}


\end{document}